---
title: "Setup"
---

Complete these steps **before** the first episode. If you run into problems, please open an
issue on the [GitHub repository]({{< var lesson.repo_url >}}/issues) or ask your instructor.

## Software

### R and RStudio

This lesson requires **R ≥ 4.2** and **RStudio ≥ 2023.06** (or another IDE of your choice).

- [Download R](https://cran.r-project.org/)
- [Download RStudio Desktop](https://posit.co/download/rstudio-desktop/)

## R packages

Install the required packages by running the following in your R console:

```{r}
#| label: install-packages
#| eval: false

# Replace with the actual packages needed for this lesson
install.packages(c(
  "tidyverse",
  "here"
  # add packages here
))
```

Verify installation:

```{r}
#| label: verify-packages
#| eval: false

# Replace with the actual packages needed for this lesson
library(tidyverse)
library(here)
# add checks here

cat("All packages loaded successfully.\n")
```

## Example data

<!-- Describe what the example dataset is and where it comes from. -->

Download the example dataset for this lesson:

```{r}
#| label: download-data
#| eval: false

# Replace with the actual data download for this lesson
download.file(
  url = "NA",   # replace with actual URL
  destfile = here("data", "example-data.csv")
)
```

::: {.callout-note}
## About the example dataset

Replace this with a brief description of the dataset used in this lesson:
what it contains, where it comes from, and any important caveats or
licensing information.
:::

## Verify your setup

Run the following chunk to confirm everything is working before Episode 1:

```{r}
#| label: verify-setup
#| eval: false

# Replace with a meaningful check for this lesson
cat("R version:", R.version$major, ".", R.version$minor, "\n", sep = "")
# add further checks as needed
```

If all checks pass, you are ready to begin [Episode 1](episodes/01-episode.qmd).
