---
title: "OBIS & Darwin Core: What They Are and Why They Matter"
description: "Meet OBIS, the global network behind it, and see why publishing your eDNA data through it matters."
time: "~60 min"
---

```{r}
#| label: setup
#| include: false
#| eval: false

library(robis)
```

::: {.callout-note icon=false}
## {{< fa bullseye >}} Learning objectives

By the end of this episode, you will be able to:

- Explain what OBIS is and how it relates to GBIF and UNESCO's Intergovernmental Oceanographic Commission (IOC).
- Describe why publishing eDNA data through a standardized system like OBIS matters, beyond simply depositing sequences in NCBI.
- Give examples of how eDNA data already published in OBIS is being reused.
- Query OBIS for existing DNA-derived occurrence records using the `robis` package.
:::

::: {.callout-important icon=false collapse=true}
## Prerequisites for this episode

- Completed [Episode 1: Introduction to eDNA & Biodiversity Data](01-introduction.qmd)
:::

---

## What is OBIS?

**OBIS — the Ocean Biodiversity Information System** — is a global, open-access data system for marine life.

- Endorsed by UNESCO's Intergovernmental Oceanographic Commission (IOC).
- Holds more than 199 million occurrence records from over 7,000 datasets.
- Data published in OBIS also flows to [GBIF](https://gbif.org) — publish once, reach both networks.
- Free to use, free to contribute to.

🌐 [obis.org](https://obis.org) · 📖 [manual.obis.org](https://manual.obis.org)

OBIS is not just a repository. It actively feeds into global biodiversity assessments, climate analyses, and conservation planning — for example the IPBES Global Assessment, UN World Ocean Assessments, and the CBD's Kunming-Montreal Global Biodiversity Framework.

### A mature, global community

OBIS is the world's largest open-access repository for marine biodiversity data, built by a genuinely global network:

- **36 nodes** across **99 countries**
- Representing **1,000+ institutions**
- And **6,000+ scientists**

OBIS sits within UNESCO's IOC (IODE, established 1961), and works alongside partners including GBIF, TDWG, EMODnet, GOOS, and DataONE. Find your regional OBIS node at [obis.org/nodes](https://obis.org/nodes) — most contributors publish through their node's Integrated Publishing Toolkit (IPT) rather than running their own infrastructure.

---

## Why publish eDNA data through OBIS?

eDNA captures biodiversity signals that other survey methods often miss:

- Cryptic and rare marine taxa
- Non-invasive, scalable sampling
- Community-level biodiversity signals

But raw sequences alone aren't enough. **Sequences deposited only in NCBI/GenBank cannot be searched by species, in time, or in space** — there's no link to coordinates or sampling dates, so they can't support distribution maps, trend analyses, or Marine Protected Area assessments.

Publishing through OBIS (and GBIF) turns a sequence into reusable biodiversity data:

| Benefit | Why it matters |
|---|---|
| Open DOI | Citable in publications |
| Long-term archive | FAIR data |
| Global reach | OBIS + GBIF |
| Sequence search | Reuse & reanalysis |

**Your sequences become findable** — OBIS's sequence search tools let others search for similar sequences across all published datasets, but only because the data is submitted in a standardized form. That standardization is the subject of the next episode.

### eDNA is already in OBIS — and being used

OBIS currently holds around 128 datasets containing eDNA-derived occurrences, totaling roughly 44.5 million records and about 2 million distinct sequences. That's only ~2% of all datasets in OBIS, but they contribute **25% of all occurrence records** — eDNA datasets tend to be large.

Real eDNA datasets already published include eDNA surveys from Monterey Bay, 16S rRNA metabarcoding of pico- to mesoplankton, and fish eDNA surveys across multiple ocean basins. This data is being used for species distribution modeling, biodiversity assessments, MPA baseline studies, and global meta-analyses.

You can query these records directly from R using the [`robis`](https://iobis.github.io/robis/) package, filtering for the DNA extension:

```{r}
#| eval: false
library(robis)

# occurrence records that include the DNA Derived Data extension
occurrence(
  hasextensions = "DNADerivedData",
  taxonid = 127160  # e.g. Actinopteri
)
```

::: {.callout-caution icon=false}
## 🧪 Exercise 2.1: Explore DNA-derived occurrences in OBIS

Using `robis::occurrence()`, retrieve DNA-derived occurrence records for a marine group you're interested in (look up its WoRMS AphiaID first, or search by `scientificname` instead of `taxonid`). How many records do you get back, and which datasets do they come from?

::: {.callout-caution icon=false collapse=true}
## Solution

```{r}
#| eval: false
library(robis)

recs <- occurrence(
  scientificname = "Actinopteri",
  hasextensions = "DNADerivedData"
)

nrow(recs)
unique(recs$datasetName)
```

Try changing `scientificname` to a taxon of your own interest — the AphiaID or scientific name both work with `occurrence()`.
:::
:::

---

## Resources and community

| Resource | Link |
|---|---|
| OBIS Manual (DNA chapter) | [manual.obis.org/dna_data.html](https://manual.obis.org/dna_data.html) |
| GBIF DNA publishing guide | [doi.org/10.35035/DOC-VF1A-NR22](https://doi.org/10.35035/DOC-VF1A-NR22) |
| OBIS helpdesk | helpdesk@obis.org |

The OBIS community is genuinely welcoming to beginners, and the helpdesk responds quickly if you get stuck formatting your own dataset.

---

::: {.instructor-only}
::: {.callout-warning icon=false}
## {{< fa chalkboard-teacher >}} Instructor notes

**Estimated time:** 60 minutes

**Pacing notes:**

- Spend more time on "why publish through OBIS" than on OBIS trivia — the NCBI-is-not-enough argument (no space/time link) is the point learners most need to internalize before the standards episode.
- If running Exercise 2.1 live, warn learners that `robis::occurrence()` requires an internet connection and can be slow for very common taxa; suggest picking a narrower group.

**Common errors:**

- Confusing an AphiaID with a GBIF taxon key — `robis` wants the AphiaID (or a plain scientific name).
- Expecting `occurrence()` results to already contain sequence data — the DNA sequences themselves live in the DNADerivedData extension, fetched separately with `occurrence(..., extensions = "DNADerivedData")` if needed.

**Discussion prompts:**

- What biodiversity questions could *not* be answered if eDNA data stayed only in NCBI/GenBank?
- Why might publishing through a regional OBIS node be easier than running your own IPT?
:::
:::

---

::: {.callout-tip icon=false}
## {{< fa key >}} Key points

- OBIS is a global, UNESCO IOC-endorsed system for marine biodiversity data that also feeds data to GBIF; you publish once and reach both.
- Raw sequences in NCBI alone can't answer *where* or *when* a species was detected — publishing through OBIS links sequences to space, time, and method.
- eDNA datasets are a small fraction of OBIS's holdings by dataset count, but a large fraction of its records, and are already used in real biodiversity assessments.
- `robis::occurrence(hasextensions = "DNADerivedData")` lets you query existing DNA-derived occurrence records directly from R.
:::
